### 2020-08-31 thesis/appendix_1.tex
same with bear
small fixes

%\begin{landscape}
\begin{longtable}[c]{|p{5cm}|p{11cm}|}
%\centering
\caption{Description of the 5 umbrella species (adapted from \cite{rayfield_priorisation_2018}).}
\label{tab:species} \\
%\begin{tabular}{|p{5cm}|p{10cm}|}
\hline
\hline
\textbf{Species} & \textbf{Description} \\ \hline
Northern short-tailed Shrew \newline \textit{Blarina brevicauda} & Abundant small fossorial mammal. This highly active species can live in a diversity of habitats (grasslands, old fields, marshy areas, gardens, and some developed areas) but is mainly found in deciduous and mixed old forests with thick understories that provide good cover for hiding from predators. It feeds primarily on earthworms found in areas with moist soils. It has a high reproductive rate and is generally a poor disperser although it can cross gaps of 50-100m. \\ \hline
American marten \newline \textit{Martes americana} & Small vagile carnivorous predator. Found in core areas of dense (\textgreater{}60\% cover) and old (\textgreater{}70 yr) coniferous or mixed forests with complex vertical and horizontal structure. It requires large home ranges (above hundreds of ha). It generally avoids large openings and clearings (above few hundred meters wide) but crosses roads and frozen rivers easily. Deep persistent snow pack is a habitat critical element as it excludes predators (Canis latrans) and competitors (Martes pennanti) and provides good hunting conditions. This forest specialist is particularly sensitive to human activities. Juveniles are able to cover tens of kilometers when dispersing, more than what would be expected from body mass-based estimates. Trapped for its fur, this species has patrimonial and economical importance. \\ \hline
Red-Backed Salamander \newline \textit{Plethodon cinereus} & Terrestrial salamander. This sedentary and territorial forest-dwelling species lives under the leaf litter or coarse woody debris in mature and moist deciduous and mixed forests. It is a poor disperser that uses tens of square meters as a home range and rarely ventures more than 50 m in open fields. Roads and edges (up to 20-30 m) have a negative effect on populations’ densities and reduce individual movements. \\ \hline
Wood Frog \newline \textit{Rana sylvatica} & Forest-specialist amphibian. This species prefers mixed and coniferous stands with closed canopy (\textgreater 40\%) and moist soil covered with woody debris (for egg deposition) but can adapt to other closed habitats. Both aquatic (palustrine, fish-free wetlands, not open and permanent ones) and terrestrial habitats are essential, and distance between both should not exceed ca. 600 m. It is sensitive to forest edge (25-35 m), gaps, high intensity agriculture, human developments and recent clearcuts that can act as barriers to movement. This poor disperser is particularly sensitive to roads and habitat loss (fidelity to first breeding pond). \\ \hline
Black Bear \newline \textit{Ursus americanus} & Large opportunistic omnivorous mammal. This species likes deciduous and mixed mature forest with dense cover interspersed with small clearings and early-successional stages of forest that are rich in berry production (depends on main soil surface deposit). It uses broad territories (tens to hundreds of square km) to follow fruiting season by going upslope with the season. It is an effective seed disperser because it can travel long distances (up to 390 km), in particular male juveniles. It clearly avoids human activity (up to 5 km) and roads (up to 800 m), in particular highways, and is likely to take a detour instead of crossing a 60 m gap. \\
\hline
\hline
%\end{tabular}
\end{longtable}
%\end{landscape}

@@

\begin{table}[h!]
\centering
\caption{Summary table of the community workshop: opportunities and challenges (non spatial elements).}
\label{tab:opp_chall_ns}
\begin{tabular}{m{0.15\textwidth}lm{0.5\textwidth}l}
\hline
\textbf{Table} &
  \textbf{\begin{tabular}[c]{@{}l@{}}Atout/Contrainte \\ (Opportunity/Challenge)\end{tabular}} &
  \textbf{\begin{tabular}[c]{@{}l@{}}Contenu\\ (Content)\end{tabular}} &
  \textbf{Score} \\ \hline
\multirow{5}{*}{Centre} &
  \multirow{3}{*}{Atouts} &
  Présence d'organisme environnementaux &
  8.00 \\ \cline{3-4} 
                       &                              & Basses terres: lien prioritaire l’échelle nationale & 4.00  \\ \cline{3-4} 
                       &                              & Programme ALUS                                      & 1.00  \\ \cline{2-4} 
                       & \multirow{2}{*}{Contraintes} & CPTAQ                                               & 18.00 \\ \cline{3-4} 
                       &                              & Besoin de rentabilité des entreprises agricoles     & 5.00  \\ \hline
\multirow{3}{*}{Est}   & Atouts                       & Vocation forestiere existante                       & 8.00  \\ \cline{2-4} 
                       & \multirow{2}{*}{Contraintes} & Tenure privée des terres                            & 2.00  \\ \cline{3-4} 
                       &                              & Plusieurs territoires couverts                      & 1.00  \\ \hline
\multirow{5}{*}{Nord} &
  \multirow{2}{*}{Atouts} &
  Réglementation favorable maintien couvert bois et corridor métropolitain &
  16.20 \\ \cline{3-4} 
                       &                              & PRMHH                                               & 10.80 \\ \cline{2-4} 
                       & \multirow{3}{*}{Contraintes} & Usage agricole prédominant                          & 44.33 \\ \cline{3-4} 
                       &                              & Compréhension et participation citoyenne            & 14.40 \\ \cline{3-4} 
                       &                              & Coûts pour faire de la connectivité                 & 10.80 \\ \hline
\multirow{6}{*}{Ouest} & \multirow{2}{*}{Atouts}      & Routes cours d'eau                                  & 3.00  \\ \cline{3-4} 
                       &                              & Presence de sols pauvre                             & 3.00  \\ \cline{2-4} 
                       & \multirow{4}{*}{Contraintes} & Les réalités économiques                            & 50.00 \\ \cline{3-4} 
                       &                              & Terres privées                                      & 12.00 \\ \cline{3-4} 
                       &                              & LPTAA                                               & 7.20  \\ \cline{3-4} 
                       &                              & Le monde politique                                  & 4.00  \\ \hline
\multirow{6}{*}{Montérégie} &
  \multirow{4}{*}{Atouts} &
  Milieu agricole: potentiel de faire des corridors si un levier est trouvé &
  12.60 \\ \cline{3-4} 
                       &                              & Présence d'acteurs locaux et education              & 8.00  \\ \cline{3-4} 
                       &                              & PRMHH                                               & 3.00  \\ \cline{3-4} 
                       &                              & Permettre l'aménagement forestier                   & 3.00  \\ \cline{2-4} 
                       & \multirow{2}{*}{Contraintes} & Routes et autoroutes                                & 4.00  \\ \cline{3-4} 
                       &                              & Terres privées, présence prédominantes              & 3.00  \\ \hline
\end{tabular}
\end{table}

@@

% Radar All
\begin{figure}[h!]
\makebox[\textwidth]{
  \includegraphics[width=1.3\textwidth]{thesis/figures/radar_ggradar_both.png}
}
 \caption{Change in mean flow (in \% of the the 2010 flow) between 2010 (observed) and 2100 (predicted), contrasting BAU scenario (solid line) with all other scenarios from borth chapters - radar graph.}
 \label{fig:flow_radar_both}
\end{figure}

>>>>

%\begin{landscape}
\begin{longtable}[c]{|p{5cm}|p{11cm}|}
%\centering
\caption{Description of the 5 umbrella species (adapted from \cite{rayfield_priorisation_2018}).}
\label{tab:species} \\
%\begin{tabular}{|p{5cm}|p{10cm}|}
\hline
\hline
\textbf{Species} & \textbf{Description} \\ \hline
Northern short-tailed Shrew \newline \textit{Blarina brevicauda} & Abundant small fossorial mammal. This highly active species can live in a diversity of habitats (grasslands, old fields, marshy areas, gardens, and some developed areas) but is mainly found in deciduous and mixed old forests with thick understories that provide good cover for hiding from predators. It feeds primarily on earthworms found in areas with moist soils. It has a high reproductive rate and is generally a poor disperser although it can cross gaps of 50-100m. \\ \hline
American marten \newline \textit{Martes americana} & Small vagile carnivorous predator. Found in core areas of dense (\textgreater{}60\% cover) and old (\textgreater{}70 yr) coniferous or mixed forests with complex vertical and horizontal structure. It requires large home ranges (above hundreds of ha). It generally avoids large openings and clearings (above a few hundred meters wide) but crosses roads and frozen rivers easily. Deep persistent snow pack is a habitat critical element as it excludes predators (Canis latrans) and competitors (Martes pennanti) and provides good hunting conditions. This forest specialist is particularly sensitive to human activities. Juveniles are able to cover tens of kilometers when dispersing, more than what would be expected from body mass-based estimates. Trapped for its fur, this species has patrimonial and economical importance. \\ \hline
Red-Backed Salamander \newline \textit{Plethodon cinereus} & Terrestrial salamander. This sedentary and territorial forest-dwelling species lives under the leaf litter or coarse woody debris in mature and moist deciduous and mixed forests. It is a poor disperser that uses tens of square meters as a home range and rarely ventures more than 50 m in open fields. Roads and edges (up to 20-30 m) have a negative effect on populations’ densities and reduce individual movements. \\ \hline
Wood Frog \newline \textit{Rana sylvatica} & Forest-specialist amphibian. This species prefers mixed and coniferous stands with closed canopy (\textgreater 40\%) and moist soil covered with woody debris (for egg deposition) but can adapt to other closed habitats. Both aquatic (palustrine, fish-free wetlands, not open and permanent ones) and terrestrial habitats are essential, and distance between both should not exceed ca. 600 m. It is sensitive to forest edge (25-35 m), gaps, high intensity agriculture, human developments and recent clearcuts that can act as barriers to movement. This poor disperser is particularly sensitive to roads and habitat loss (fidelity to the first breeding pond). \\ \hline
Black Bear \newline \textit{Ursus americanus} & Large opportunistic omnivorous mammal. This species likes deciduous and mixed mature forest with dense cover interspersed with small clearings and early-successional stages of forest that are rich in berry production (depends on main soil surface deposit). It uses broad territories (tens to hundreds of square km) to follow the fruiting season by going upslope with the season. It is an effective seed disperser because it can travel long distances (up to 390 km), in particular male juveniles. It clearly avoids human activity (up to 5 km) and roads (up to 800 m), in particular highways, and is likely to take a detour instead of crossing a 60 m gap. \\
\hline
\hline
%\end{tabular}
\end{longtable}
%\end{landscape}

@@

\begin{table}[h!]
\centering
\caption{Summary table of the community workshop: opportunities and challenges (non spatial elements).}
\label{tab:opp_chall_ns}
\begin{tabular}{m{0.15\textwidth}lm{0.5\textwidth}l}
\hline
\textbf{Table} &
  \textbf{\begin{tabular}[c]{@{}l@{}}Atout/Contrainte \\ (Opportunity/Challenge)\end{tabular}} &
  \textbf{\begin{tabular}[c]{@{}l@{}}Contenu\\ (Content)\end{tabular}} &
  \textbf{Score} \\ \hline
\multirow{5}{*}{Centre} &
  \multirow{3}{*}{Atouts} &
  Présence d'organisme environnementaux &
  8.00 \\ \cline{3-4} 
                       &                              & Basses terres: lien prioritaire l’échelle nationale & 4.00  \\ \cline{3-4} 
                       &                              & Programme ALUS                                      & 1.00  \\ \cline{2-4} 
                       & \multirow{2}{*}{Contraintes} & CPTAQ                                               & 18.00 \\ \cline{3-4} 
                       &                              & Besoin de rentabilité des entreprises agricoles     & 5.00  \\ \hline
\multirow{3}{*}{Est}   & Atouts                       & Vocation forestiere existante                       & 8.00  \\ \cline{2-4} 
                       & \multirow{2}{*}{Contraintes} & Tenure privée des terres                            & 2.00  \\ \cline{3-4} 
                       &                              & Plusieurs territoires couverts                      & 1.00  \\ \hline
\multirow{5}{*}{Nord} &
  \multirow{2}{*}{Atouts} &
  Réglementation favorable maintien couvert bois et corridor métropolitain &
  16.20 \\ \cline{3-4} 
                       &                              & PRMHH                                               & 10.80 \\ \cline{2-4} 
                       & \multirow{3}{*}{Contraintes} & Usage agricole prédominant                          & 44.33 \\ \cline{3-4} 
                       &                              & Compréhension et participation citoyenne            & 14.40 \\ \cline{3-4} 
                       &                              & Coûts pour faire de la connectivité                 & 10.80 \\ \hline
\multirow{6}{*}{Ouest} & \multirow{2}{*}{Atouts}      & Routes cours d'eau                                  & 3.00  \\ \cline{3-4} 
                       &                              & Presence de sols pauvre                             & 3.00  \\ \cline{2-4} 
                       & \multirow{4}{*}{Contraintes} & Les réalités économiques                            & 50.00 \\ \cline{3-4} 
                       &                              & Terres privées                                      & 12.00 \\ \cline{3-4} 
                       &                              & LPTAA                                               & 7.20  \\ \cline{3-4} 
                       &                              & Le monde politique                                  & 4.00  \\ \hline
\multirow{6}{*}{Montérégie} &
  \multirow{4}{*}{Atouts} &
  Milieu agricole: potentiel de faire des corridors si un levier est trouvé &
  12.60 \\ \cline{3-4} 
                       &                              & Présence d'acteurs locaux et éducation              & 8.00  \\ \cline{3-4} 
                       &                              & PRMHH                                               & 3.00  \\ \cline{3-4} 
                       &                              & Permettre l'aménagement forestier                   & 3.00  \\ \cline{2-4} 
                       & \multirow{2}{*}{Contraintes} & Routes et autoroutes                                & 4.00  \\ \cline{3-4} 
                       &                              & Terres privées, présence prédominantes              & 3.00  \\ \hline
\end{tabular}
\end{table}

@@

% Radar All
\begin{figure}[h!]
\makebox[\textwidth]{
  \includegraphics[width=1.3\textwidth]{thesis/figures/radar_ggradar_both.png}
}
 \caption{Change in mean flow (in \% of the 2010 flow) between 2010 (observed) and 2100 (predicted), contrasting BAU scenario (solid line) with all other scenarios from borth chapters - radar graph.}
 \label{fig:flow_radar_both}
\end{figure}

### 2020-08-31 thesis/chap_1.tex
small fixes

% [CC as an adaptation measure, but CC designs are vulnerable]
Connectivity conservation has often been identified out as a method for adaptation to climate and land use change \citep{costanza_landscape_2019}. One recurring argument is that preserving landscape connectivity could be critical for ensuring that species can shift their ranges \citep{krosby_ecological_2010, keeley_making_2018} in response to climate change. Another argument is that it is more important than ever to connect habitats in landscapes increasingly hostile to species movement \citep{ellis_anthropogenic_2010}. These arguments make the case for the design of Connected Protected Area Networks (CPAN, \cite{opdam_ecological_2006, stewart_corridors_2019, gonzalez_spatial_2017}). Those networks are meant to allow the movement of species across multiple scales, which may require them to include both natural and semi-natural habitat, with different levels of protections \citep{daloia_coupled_2019}. This heterogeneity is likely to be reinforced by the multiplicity of urban and land use planning practices that exist on the landscape, making CPAN designs particularly vulnerable to changes in land use. It is therefore important to test which specific CPAN designs would be most resilient to future  change.  One possible approach to exploring this resilience is to use landscape-level models that project the future state of land-cover.

@@

The first step consisted in reclassifying the AAFC land use dataset into 4 categories: the 3 categories of our land use change model plus roads and wetlands. Then we aggregated the rasters by a factor of 3 (using the mode of the neighborhood, changing the resolution to 90m). We applied this aggregation to all rasters (for continous data we used the mean and not the mode). This allowed to reduce run time as well as memory requirements. Roads were removed during aggregation: this was necessary as the process tends to break them up, creating holes in them. Major roads from the Adresses Quebec dataset were rasterized and patched on top of the resulting maps. Then, all the pixels classified as forest were reclassified as null, and the resulting maps were patched with the forest cover from the LANDIS simulations' starting conditions (state of forest types in 2010). Some cells remained null as the forest cover in 1990 was larger than that of 2010. We filled those NULL values using modal imputation under two different neighbourhood window sizes (13 and 25 pixels). Next, a 5km buffer was added around Montérégie so as to improve the Circuitscape results later down the pipeline (see the Connectivity Analysis section). Not all NA cells were due to the difference in forest cover between 1990 and 2010, some were due to the fact that the LANDIS run did not cover some parts of the 5km buffer. Those cells were filled with the value corresponding to the forest type Mixed Forest of Medium Age. We decided on this imputed value because it is among the most suitable for all species and therefore unlikely to cause significant changes in the Circuitscape runs, even if it does not correspond to the mode or the mean for the region. In addition, because no change takes place in the buffer, those values can be equivalent to a "padding" around the landscape.

@@

The ST-Sim model uses three states: Forest, Urban and Agriculture. Forest was further subdivided into classes of age and type, derived from the LANDIS outputs: three forest types (\textbf{decididuous}, \textbf{mixt}, and \textbf{coniferous}) and three age classes (\textbf{young}, \textbf{medium}, and \textbf{old}) for a total of 9 classes.
The model allows for 4 groups of transitions:
\begin{itemize}
\item{\textbf{Urbanisation}}: groups \textbf{deforestation} (Forest $\Rightarrow$ Urban) and \textbf{agricultural loss} (Agriculture $\Rightarrow$ Urban).
\item{\textbf{Agricultural expansion}} (Forest $\Rightarrow$ Urban).
\item{\textbf{Forest Internals}}: groups the 72 combinations of forest change between the 9 forest internal states.
\end{itemize}
The model takes in 3 different strata. A stratum in ST-Sim refers to a subset of the entire landscape being considered for analysis. Strata are typically used to divide the overall landscape into regions or zones, each of which can have different model parameters specified. The three stratum are:
\begin{itemize}
\item{Primary stratum:} delineates Montérégie from the  surrounding buffer and highlights protected areas.
\item{Secondary stratum:} delineates municipalities of the region, to allow to set targets specific to each municipality.
\item{Tertiary stratum:} delineates LANDIS land types, to allow to set transition multipliers for forest types that match the LANDIS simulations. \\
\end{itemize}

@@

In an attempt to go beyond the simple analysis of current flow value, we reproduced the workflow for pinch point analysis laid out in \cite{pelletier_forest_2017}. In their analysis of connectivity at the scale of Canada, using Circuitscape \citeauthor{pelletier_forest_2017} used the SURF algorithm (Speedep Up Robust Features) to detect narrow regions of high contrast in the flow surface (or "pinch points", as described in \cite{mcrae_using_2008}). SURF is a computer vision algorithm used to reliably detect and match features in images, regardless of their orientation (see \cite{leonardis_surf_2006} for an in-depth description of the algorithm). We attempted to reproduce the results of \citeauthor{pelletier_forest_2017}, although no code was publically available. We use the OpenCV4 python Library implementation of the SURF software.

In order to get to a certain level of reliability in the results across iterations, it was found that it was important to equalize the histograms of pixel flow values. With this modification, we allow for areas of lower local contrast to gain a higher contrast. This is achieved by taking most frequent intensity values and spreading them out. This method usually increases the global contrast of images \citep{garg_comparative_2017}. But even with this helpful modification to the flow surfaces, it was difficult to accurately identify pinch points in the images. However, the method still served to identify and detect relevant flow features in the images. In their implementation of the method, \citeauthor{pelletier_forest_2017} mention that SURF could potentially be used to measure of "flow complexity". Therefore, we hope this step of the analysis can serve as a way to measure the changes in complexity of the flow that arise under different scenarios. We recognize the limitations of this method at this stage but it does provide an interesting use case for the application of computer vision methods to the interpretation of mapped results from connectivity analyses.

@@

The clustering and subsequent ordination of land use change profiles showed that Montérégie has 4 different transition profiles (see figures \ref{fig:PCAtrans}, and \ref{fig:maptrans}):
\begin{itemize}[leftmargin=0.5cm]
 \item \textbf{Urban Spread / Deforestation}: forest fragmentation is progressing mainly via the growth of urban land (in the west) or development for holiday cottages (“villegiatives pressures”, in the east).
 \item \textbf{Urban Spread / Agricultural loss}: agriculture is losing ground to urban land.
 \item \textbf{Agricultural Expansion / Fragmentation}: forest is being lost toagriculture in those municipalities where forest is still quite present.
 \item \textbf{Agricultural Expansion / Deforestation}: forest is already scarce and is being replaced by agricultural land.
These results show marked regional trends with a front line of fragmentation and deforestation on both sides of the region and along the Richelieu river. \\ % (Fig. \ref{fig:map}).
\end{itemize}

@@

\begin{figure}[h!]
\makebox[\textwidth]{
  \includegraphics[width=1.3\textwidth]{thesis/figures/connectivity_decrease_x5species_historic.png}
}
 \caption{Change in mean flow (in \% of the the 1990 flow) between 1990 and 2010, contrasting observed historic change (1990-2010) and simulated change from historic model run.}
 \label{fig:flow_historic}
\end{figure}

@@

\begin{figure}[h!]
\makebox[\textwidth]{
  \includegraphics[width=1.3\textwidth]{thesis/figures/connectivity_decrease_x5species_chap1.png}
}
 \caption{Change in mean flow (in \% of the the 2010 flow) between 2010 (observed) and 2100 (predicted), contrasting BAU scenario (solid line) with other land use change scenarios - linear graph.}
 \label{fig:flow_linear_1}
\end{figure}

@@

\begin{figure}[h!]
\makebox[\textwidth]{
  \includegraphics[width=1.3\textwidth]{thesis/figures/radar_ggradar_chap1.png}
}
 \caption{Change in mean flow (in \% of the the 2010 flow) between 2010 (observed) and 2100 (predicted), contrasting BAU scenario with other land use change scenarios - radar graph.}
 \label{fig:flow_radar_1}
\end{figure}

@@

\begin{figure}[h!]
\makebox[\textwidth]{
  \includegraphics[width=1.3\textwidth]{thesis/figures/surf_chap1.png}
}
 \caption{Change in the number of features (in \% of the the 2010 flow) identified by the SURF analyssis between 2010 (observed) and 2100 (predicted), contrasting BAU scenario with other land use change scenarios - linear graph.}
 \label{fig:surf_linear_1}
\end{figure}

@@

\begin{figure}[h!]
\makebox[\textwidth]{
  \includegraphics[width=1.3\textwidth]{thesis/figures/surf_radar_chap1.png}
}
 \caption{Change in the number of features (in \% of the the 2010 flow) identified by the SURF analyssis between 2010 (observed) and 2100 (predicted), contrasting BAU scenario with other land use change scenarios - radar graph.}
 \label{fig:surf_radar_1}
\end{figure}

>>>>

% [CC as an adaptation measure, but CC designs are vulnerable]
Connectivity conservation has often been identified as a method for adaptation to climate and land use change \citep{costanza_landscape_2019}. One recurring argument is that preserving landscape connectivity could be critical for ensuring that species can shift their ranges \citep{krosby_ecological_2010, keeley_making_2018} in response to climate change. Another argument is that it is more important than ever to connect habitats in landscapes increasingly hostile to species movement \citep{ellis_anthropogenic_2010}. These arguments make the case for the design of Connected Protected Area Networks (CPAN, \cite{opdam_ecological_2006, stewart_corridors_2019, gonzalez_spatial_2017}). Those networks are meant to allow the movement of species across multiple scales, which may require them to include both natural and semi-natural habitat, with different levels of protections \citep{daloia_coupled_2019}. This heterogeneity is likely to be reinforced by the multiplicity of urban and land use planning practices that exist on the landscape, making CPAN designs particularly vulnerable to changes in land use. It is therefore important to test which specific CPAN designs would be most resilient to future  change.  One possible approach to exploring this resilience is to use landscape-level models that project the future state of land-cover.

@@

The first step consisted in reclassifying the AAFC land use dataset into 4 categories: the 3 categories of our land use change model plus roads and wetlands. Then we aggregated the rasters by a factor of 3 (using the mode of the neighborhood, changing the resolution to 90m). We applied this aggregation to all rasters (for continuous data we used the mean and not the mode). This allowed us to reduce run time as well as memory requirements. Roads were removed during aggregation: this was necessary as the process tends to break them up, creating holes in them. Major roads from the Adresses Quebec dataset were rasterized and patched on top of the resulting maps. Then, all the pixels classified as forest were reclassified as null, and the resulting maps were patched with the forest cover from the LANDIS simulations' starting conditions (state of forest types in 2010). Some cells remained null as the forest cover in 1990 was larger than that of 2010. We filled those NULL values using modal imputation under two different neighbourhood window sizes (13 and 25 pixels). Next, a 5km buffer was added around Montérégie so as to improve the Circuitscape results later down the pipeline (see the Connectivity Analysis section). Not all NA cells were due to the difference in forest cover between 1990 and 2010, some were due to the fact that the LANDIS run did not cover some parts of the 5km buffer. Those cells were filled with the value corresponding to the forest type Mixed Forest of Medium Age. We decided on this imputed value because it is among the most suitable for all species and therefore unlikely to cause significant changes in the Circuitscape runs, even if it does not correspond to the mode or the mean for the region. In addition, because no change takes place in the buffer, those values can be equivalent to a "padding" around the landscape.

@@

The ST-Sim model uses three states: Forest, Urban and Agriculture. Forest was further subdivided into classes of age and type, derived from the LANDIS outputs: three forest types (\textbf{deciduous}, \textbf{mixt}, and \textbf{coniferous}) and three age classes (\textbf{young}, \textbf{medium}, and \textbf{old}) for a total of 9 classes.
The model allows for 4 groups of transitions:
\begin{itemize}
\item{\textbf{Urbanisation}}: groups \textbf{deforestation} (Forest $\Rightarrow$ Urban) and \textbf{agricultural loss} (Agriculture $\Rightarrow$ Urban).
\item{\textbf{Agricultural expansion}} (Forest $\Rightarrow$ Urban).
\item{\textbf{Forest Internals}}: groups the 72 combinations of forest change between the 9 forest internal states.
\end{itemize}
The model takes in 3 different strata. A stratum in ST-Sim refers to a subset of the entire landscape being considered for analysis. Strata are typically used to divide the overall landscape into regions or zones, each of which can have different model parameters specified. The three stratum are:
\begin{itemize}
\item{Primary stratum:} delineates Montérégie from the  surrounding buffer and highlights protected areas.
\item{Secondary stratum:} delineates municipalities of the region, to allow to set targets specific to each municipality.
\item{Tertiary stratum:} delineates LANDIS land types, to allow to set transition multipliers for forest types that match the LANDIS simulations. \\
\end{itemize}

@@

In an attempt to go beyond the simple analysis of current flow value, we reproduced the workflow for pinch point analysis laid out in \cite{pelletier_forest_2017}. In their analysis of connectivity at the scale of Canada, using Circuitscape \citeauthor{pelletier_forest_2017} used the SURF algorithm (Speeded Up Robust Features) to detect narrow regions of high contrast in the flow surface (or "pinch points", as described in \cite{mcrae_using_2008}). SURF is a computer vision algorithm used to reliably detect and match features in images, regardless of their orientation (see \cite{leonardis_surf_2006} for an in-depth description of the algorithm). We attempted to reproduce the results of \citeauthor{pelletier_forest_2017}, although no code was publically available. We use the OpenCV4 python Library implementation of the SURF software.

In order to get to a certain level of reliability in the results across iterations, it was found that it was important to equalize the histograms of pixel flow values. With this modification, we allow for areas of lower local contrast to gain a higher contrast. This is achieved by taking most frequent intensity values and spreading them out. This method usually increases the global contrast of images \citep{garg_comparative_2017}. But even with this helpful modification to the flow surfaces, it was difficult to accurately identify pinch points in the images. However, the method still served to identify and detect relevant flow features in the images. In their implementation of the method, \citeauthor{pelletier_forest_2017} mention that SURF could potentially be used to measure "flow complexity". Therefore, we hope this step of the analysis can serve as a way to measure the changes in complexity of the flow that arise under different scenarios. We recognize the limitations of this method at this stage but it does provide an interesting use case for the application of computer vision methods to the interpretation of mapped results from connectivity analyses.

@@

The clustering and subsequent ordination of land use change profiles showed that Montérégie has 4 different transition profiles (see figures \ref{fig:PCAtrans}, and \ref{fig:maptrans}):
\begin{itemize}[leftmargin=0.5cm]
 \item \textbf{Urban Spread / Deforestation}: forest fragmentation is progressing mainly via the growth of urban land (in the west) or development for holiday cottages (“villegiatives pressures”, in the east).
 \item \textbf{Urban Spread / Agricultural loss}: agriculture is losing ground to urban land.
 \item \textbf{Agricultural Expansion / Fragmentation}: forest is being lost to agriculture in those municipalities where forest is still quite present.
 \item \textbf{Agricultural Expansion / Deforestation}: forest is already scarce and is being replaced by agricultural land.
These results show marked regional trends with a front line of fragmentation and deforestation on both sides of the region and along the Richelieu river. \\ % (Fig. \ref{fig:map}).
\end{itemize}

@@

\begin{figure}[h!]
\makebox[\textwidth]{
  \includegraphics[width=1.3\textwidth]{thesis/figures/connectivity_decrease_x5species_historic.png}
}
 \caption{Change in mean flow (in \% of the 1990 flow) between 1990 and 2010, contrasting observed historic change (1990-2010) and simulated change from historic model run.}
 \label{fig:flow_historic}
\end{figure}

@@

\begin{figure}[h!]
\makebox[\textwidth]{
  \includegraphics[width=1.3\textwidth]{thesis/figures/connectivity_decrease_x5species_chap1.png}
}
 \caption{Change in mean flow (in \% of the 2010 flow) between 2010 (observed) and 2100 (predicted), contrasting BAU scenario (solid line) with other land use change scenarios - linear graph.}
 \label{fig:flow_linear_1}
\end{figure}

@@

\begin{figure}[h!]
\makebox[\textwidth]{
  \includegraphics[width=1.3\textwidth]{thesis/figures/radar_ggradar_chap1.png}
}
 \caption{Change in mean flow (in \% of the 2010 flow) between 2010 (observed) and 2100 (predicted), contrasting BAU scenario with other land use change scenarios - radar graph.}
 \label{fig:flow_radar_1}
\end{figure}

@@

\begin{figure}[h!]
\makebox[\textwidth]{
  \includegraphics[width=1.3\textwidth]{thesis/figures/surf_chap1.png}
}
 \caption{Change in the number of features (in \% of the 2010 flow) identified by the SURF analysis between 2010 (observed) and 2100 (predicted), contrasting BAU scenario with other land use change scenarios - linear graph.}
 \label{fig:surf_linear_1}
\end{figure}

@@

\begin{figure}[h!]
\makebox[\textwidth]{
  \includegraphics[width=1.3\textwidth]{thesis/figures/surf_radar_chap1.png}
}
 \caption{Change in the number of features (in \% of the 2010 flow) identified by the SURF analysis between 2010 (observed) and 2100 (predicted), contrasting BAU scenario with other land use change scenarios - radar graph.}
 \label{fig:surf_radar_1}
\end{figure}
